On Similarity

### 2016-05-04 paper.tex
an argument agains global similarity

Similarity is the degree of resemblance between two objects or events \cite{WCS:WCS1282}. It can be implemented as a Contrast Model where the similarity between objects is expressed as a function of their common and distinctive features \cite{Tversky1977}. The datasets provide \emph{global} similarity (relatedness) scores between word pairs. Global similarity refers to an overall comparison \cite{hahn1997concepts} and is restricted with regard to the following observations:
\begin{compactitem}
    % \item Similarity scores group words to categories that are formed around \emph{focal points}, the terms that represent clusters.
    % To detect a focal point, one needs similarities between all words, or pairings between the focal point and the words of a category for each category.
    \item Similarity is measured with respect to something: \textit{apple} is similar to \textit{orange} because both are companies or kinds of fruits, but dissimilar because apples grow where oranges do not \cite{hahn1997concepts}.
    \item Similarity scores presented by the datasets are ambiguous as it is not really clear what low similarity values mean: incompatible notions (\textit{trick} and \textit{size}) or contrast in meaning (\textit{smart} and \textit{dumb}). For example, SimLex-999 assigns low similarity scores to the pairs, 0.48 and 0.55 out of 10 respectively, but \textit{smart} and \textit{dumb} have relatively much more in common than \textit{trick} and \textit{size}!

>>>>

Similarity is the degree of resemblance between two objects or events \cite{WCS:WCS1282}. It can be implemented as a Contrast Model where the similarity between objects is expressed as a function of their common and distinctive features \cite{Tversky1977}. The datasets provide \emph{global} similarity (relatedness) scores between word pairs. Global similarity refers to an overall comparison \cite{hahn1997concepts} and is restricted with regard to the following observations:
\begin{compactitem}
    % \item Similarity scores group words to categories that are formed around \emph{focal points}, the terms that represent clusters.
    % To detect a focal point, one needs similarities between all words, or pairings between the focal point and the words of a category for each category.
    % TODO: the point should be against gathering global similarity, as most probably people will think of a respect when they give a similarity judgement.
    \item ``[S]imilarity\ldots is a relationship between two objects \emph{under a given description}'' \cite{WCS:WCS1282}. Global similarity judgement omit the description and possibly give imprecise global similarity. Is \textit{apple}'s similarity to \textit{orange} as a company the same as a fruit?
    \item Similarity scores presented by the datasets are ambiguous as it is not really clear what low similarity values mean: incompatible notions (\textit{trick} and \textit{size}) or contrast in meaning (\textit{smart} and \textit{dumb}). For example, SimLex-999 assigns low similarity scores to the pairs, 0.48 and 0.55 out of 10 respectively, but \textit{smart} and \textit{dumb} have relatively much more in common than \textit{trick} and \textit{size}!
